Assessing exposure to Atlantic Basin tropical storms in United States counties

### 2016-08-04 writing/DraftExposurePaper/DraftExposurePaper.Rmd
More work on draft

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(hurricaneexposuredata)
library(hurricaneexposure)
library(knitr)
library(ggplot2)
library(ggthemes)
library(grid)
library(gridExtra)
library(plyr)
library(dplyr)
library(tidyr)
library(choroplethr)
library(RColorBrewer)
library(weathermetrics)
library(lubridate)

@@

load("MWR_gauge_wind_data/frances_wind.RData")
frances_wind <- frances_wind %>%
        filter(!is.na(wind_sust)) %>%
        mutate(wind_sust = convert_wind_speed(wind_sust, "knots", "mps"))
frances_wind_mwr <- plot_mwr_compare_wind(frances_wind)
```

Compare data from `countyweather` and NLDAS-2 data for Miami-Dade county by storm:

```{r}
data("closest_dist")

# Create a new dataset for precip with lag -2 to 3 for fips 12086 only

check_dates <- dplyr::select(closest_dist, -storm_dist) %>%
        dplyr::mutate(closest_date = ymd(closest_date)) %>%
        dplyr::rename(day_0 = closest_date) %>%
        dplyr::mutate(fips = as.integer(fips),
                      day_0 = day_0 + days(0),
                      day_b1 = day_0 - days(1),
                      day_b2 = day_0 - days(2),
                      day_a1 = day_0 + days(1),
                      day_a2 = day_0 + days(2),
                      day_a3 = day_0 + days(3)) %>%
        dplyr::select(storm_id, fips, day_b2, day_b1,
                      day_0, day_a1, day_a2, day_a3) %>%
        tidyr::gather(key = lag, value = day, -storm_id, -fips) %>%
        dplyr::rename(date = day) %>%
        dplyr::filter(fips == "12086") %>%
        dplyr::mutate(fips = as.character(fips))
        #dplyr::mutate(day = as.numeric(format(day, "%Y%m%d")))

all_dates <- unique(check_dates$date)
all_fips <- unique(check_dates$fips)

#create timeseries for Dade precipitation data from the `countyweather` package

fips <- "12086"
# I've commented this out because we're saving the data, so we only need
# to run it once.
# county_timeseries(fips, percent_coverage = 0,
#                   date_min = "1988-01-01", date_max = "2011-12-31",
#                   var = "PRCP",
#                   out_directory = "dade_data/")
#                   #out_directory = "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/")

# miami_dade_dir <- "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/"
dade_dir <- "dade_data/"
dade_weather <- readRDS(paste0(dade_dir, fips, ".rds"))

dade_weather <- dade_weather$averaged %>%
        filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))

dade_weather$fips <- fips
dade_weather <- as.data.frame(dade_weather)

dade_weather <- dade_weather %>%
        dplyr::filter(date %in% all_dates) %>%
        dplyr::right_join(data.table(check_dates),
                          by = c("fips" = "fips", "date" = "date")) %>%
        dplyr::filter(!is.na(prcp)) %>%
        dplyr::arrange(storm_id, fips) %>%
        dplyr::mutate(fips = as.integer(fips)) %>%
        dplyr::mutate(fips = sprintf("%05d", fips),
                      lag = gsub("day_", "", lag),
                      lag = gsub("b", "-", lag),
                      lag = gsub("a", "", lag),
                      lag = as.numeric(lag))

lag_sum <- function(counties = NULL, county_weather, year_range = NULL,
                              distance_limit = NULL, rain_limit = NULL,
                              include_rain = FALSE, days_included = NULL,
                              output_vars = c("fips", "cw_precip")){

        closest_dist <- data.table::data.table(hurricaneexposuredata::closest_dist)

        if(!is.null(counties)){
                closest_dist <- closest_dist[get("fips") %in% counties]
        }

        if(!is.null(year_range)){
                closest_dist <- closest_dist[ , .(storm_id,
                                                  fips,
                                                  closest_date, storm_dist,
                                                  local_time,
                                                  closest_time_utc,
                                                  year = substring(get("closest_date"),
                                                                   1, 4)), ][
                                                                           get("year") %in%
                                                                                   year_range[1]:year_range[2]
                                                                           ]
        }

        if(!is.null(distance_limit)){
                closest_dist <- closest_dist[get("storm_dist") <=
                                                     distance_limit]
        }

        if(include_rain){
                df <- data.table::data.table(county_weather)
                df <- df[get("lag") %in% days_included]
                df <- df[ , .(cw_precip = sum(get("prcp"))),
                              by = .(fips, storm_id)]
                closest_dist <- merge(closest_dist, df, all.x = TRUE,
                                      by = c("storm_id", "fips"))
                if(!is.null(rain_limit)){
                        closest_dist <- closest_dist[get("cw_precip") >=
                                                             rain_limit]
                }
        }

        closest_dist <- closest_dist[ , .SD, , .SDcols = output_vars]
        return(closest_dist)
}

countyweather_rain <- function(counties, county_weather, start_year, end_year,
                        rain_limit, dist_limit,
                        days_included = c(-2, -1, 0, 1)){

        county_rain_storm_df <- lag_sum(counties = counties,
                                           county_weather = county_weather,
                                           year_range = c(start_year, end_year),
                                           distance_limit = dist_limit,
                                           rain_limit = rain_limit,
                                           include_rain = TRUE,
                                           days_included = days_included,
                                           output_vars = c("storm_id", "fips",
                                                           "closest_date",
                                                           "storm_dist",
                                                           "cw_precip",
                                                           "local_time",
                                                           "closest_time_utc"))
        rain_storm_df <- filter_storm_data(counties = counties,
                                           year_range = c(start_year, end_year),
                                           distance_limit = dist_limit,
                                           rain_limit = rain_limit,
                                           include_rain = TRUE,
                                           days_included = days_included,
                                           output_vars = c("storm_id", "fips",
                                                           "closest_date",
                                                           "storm_dist",
                                                           "tot_precip",
                                                           "local_time",
                                                           "closest_time_utc"))
        rain_df <- merge(county_rain_storm_df, rain_storm_df, all.x = TRUE,
                         by = c("fips", "storm_id", "closest_date", "storm_dist",
                                "local_time", "closest_time_utc"))
        return(rain_df)
}

miami_rain <- countyweather_rain(counties = "12086", county_weather = dade_weather,
                                 start_year = 1988, end_year = 2011,
                                 rain_limit = 0, dist_limit = 2000)
miami_plot <- ggplot(miami_rain, aes(x = cw_precip, y = tot_precip)) +
        geom_hline(aes(yintercept = 75), color = "lightgray") +
        geom_vline(aes(xintercept = 75), color = "lightgray") +
        geom_abline(aes(intercept = 0, slope = 1), color = "gray", alpha = 0.5) +
        # geom_point(aes(size = prcp_reporting), alpha = 0.2) +
        geom_point(alpha = 0.5) +
        # geom_text(aes(x = monitor_rain + 100, label = storm_id)) +
        xlim(c(0, 275)) + ylim(c(0, 275)) +
        theme_few() +
        scale_size_continuous(guide = "none") +
        xlab("Rainfall (mm) based on \naveraged county monitors") +
        ylab("Rainfall (mm) based on \nNLDAS-2 county data") +
        ggtitle("Monitor versus NLDAS rainfall estimates \nfor Miami-Dade County, FL, by storm")

miami_plot
```

```{r}
load("MWR_gauge_wind_data/ike_wind.RData")
ike_wind <- ike_wind %>%
        filter(!is.na(max_sust_kt)) %>%
        mutate(wind_sust = convert_wind_speed(max_sust_kt, "knots", "mps"))
ike_wind_mwr <- plot_mwr_compare_wind(ike_wind)

@@

Map of the count of county exposures to the top 2,000 county-storm pairs by distance-based exposure, based on "exposed" = distance of 100 kilometers or less:

```{r echo = FALSE, warning=FALSE, message=FALSE}
distance_2000 <- distance_exposure %>%
        arrange(storm_dist) %>% 
        slice(1:2000) %>% 
        group_by(fips) %>%
        summarise(count = n()) %>%
        rename(region = fips) %>%
        mutate(region = as.numeric(region), count = as.numeric(count))

distance_2000$value <- cut(distance_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
                   include.lowest = TRUE)
exposure_palette <- brewer.pal(8, name = "YlGnBu")

map_dist_2000 <- CountyChoropleth$new(distance_2000)
map_dist_2000$set_zoom(eastern_states)
map_dist_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
                                                values = exposure_palette,
                                                na.value = "white")
map_dist_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Distance Metric"
a <- map_dist_2000$render()
```

```{r echo = FALSE, warning=FALSE, message=FALSE}
# Get the rain-based exposure for all counties over the study period, 
# using a rain limit of 75 mm and a distance limit of 500 km
rain_exposure <- county_rain(counties = my_fips,
                             start_year = 1988, end_year = 2011,
                             rain_limit = 75, dist_limit = 500,
                             days_include = c(-1, 0, 1))

rain_2000 <- rain_exposure %>%
        arrange(desc(tot_precip)) %>%
        slice(1:2000) %>%
        group_by(fips) %>%
        summarise(count = n()) %>%
        rename(region = fips) %>%
        mutate(region = as.numeric(region), count = as.numeric(count))

rain_2000$value <- cut(rain_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
                   include.lowest = TRUE)
exposure_palette <- brewer.pal(8, name = "YlGnBu")

map_rain_2000 <- CountyChoropleth$new(rain_2000)
map_rain_2000$set_zoom(eastern_states)
map_rain_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
                                                values = exposure_palette,
                                                na.value = "white")
map_rain_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Rain Metric"
b <- map_rain_2000$render()
```

```{r fig.width = 6, fig.height = 8, echo = FALSE, message = FALSE, warning = FALSE}
grid.arrange(a, b, ncol = 1)
```

Map of county exposures to the top 2,000 county-storm pairs by wind-based exposure:

```{r fig.width = 6, fig.height = 8, echo = FALSE, message = FALSE, warning = FALSE}
#TS winds and above
TS_wind_exposure <- county_wind(counties = my_fips,
                             start_year = 1988,
                             end_year = 2011,
                             wind_limit = 18,
                             wind_var = "max_sust")

TS_wind_2000 <- TS_wind_exposure %>% 
        arrange(desc(max_sust)) %>%
        slice (1:2000) %>%
        group_by(fips) %>%
        summarise(count = n()) %>%
        rename(region = fips) %>%
        mutate(region = as.numeric(region), count = as.numeric(count)) # Max count of 178 is a NA value
TS_wind_2000 <- subset(TS_wind_2000, !is.na(TS_wind_2000$region)) # Remove NA observation
                                                                       # County pairs
TS_wind_2000$value <- cut(TS_wind_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
                   include.lowest = TRUE)
exposure_palette <- brewer.pal(8, name = "YlGnBu")
map_TS_2000 <- CountyChoropleth$new(TS_wind_2000)
map_TS_2000$set_zoom(eastern_states)
map_TS_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
                                                values = exposure_palette,
                                                na.value = "white")
map_TS_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor TS Wind Metric"
map_TS_2000$render()
```

>>>>

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(hurricaneexposuredata)
library(hurricaneexposure)
library(knitr)
library(ggplot2)
library(ggthemes)
library(grid)
library(gridExtra)
library(plyr)
library(dplyr)
library(tidyr)
library(choroplethr)
library(RColorBrewer)
library(weathermetrics)
library(lubridate)
library(xtable)

@@

load("MWR_gauge_wind_data/frances_wind.RData")
frances_wind <- frances_wind %>%
        filter(!is.na(wind_sust)) %>%
        mutate(wind_sust = convert_wind_speed(wind_sust, "knots", "mps"))
frances_wind_mwr <- plot_mwr_compare_wind(frances_wind)

load("MWR_gauge_wind_data/ike_wind.RData")
ike_wind <- ike_wind %>%
        filter(!is.na(max_sust_kt)) %>%
        mutate(wind_sust = convert_wind_speed(max_sust_kt, "knots", "mps"))
ike_wind_mwr <- plot_mwr_compare_wind(ike_wind)

@@

We first investigated whether different exposure metrics resulted in similar numbers of counties classified as "exposed" to a storm (Figure \ref{fig:county_number_cor}, Table \ref{tab:county_number}).

```{r county_number_cor, echo = FALSE, fig.width = 8, fig.height = 8, fig.align = "center", fig.cap = "Association between the number of counties classified as `exposed` for a storm using different exposure metrics. Each small plot shows this association, with each point representing a storm, with positions on the x- and y-axes showing the number of counties classified as `exposed` for that storm based on one of the exposure metrics considered in this study. Study storms from 1996 to 2011 are included in each plot."}
dist_exp <- distance_exposure %>%
        select(storm_id, fips) %>%
        mutate(dist_exposed = 1)
rain_exp <- rain_exposure %>%
        select(storm_id, fips) %>%
        mutate(rain_exposed = 1)
wind_exp <- wind_exposure %>%
        select(storm_id, fips) %>%
        mutate(wind_exposed = 1)
flood_exp <- flood_exposure %>%
        select(storm_id, fips) %>%
        mutate(flood_exposed = 1,
               storm_id = as.character(storm_id))
tornado_exp <- tornado_exposure %>%
        select(storm_id, fips) %>%
        mutate(tornado_exposed = 1,
               storm_id = as.character(storm_id))

all_exp <- dist_exp %>%
        full_join(rain_exp, by = c("storm_id", "fips")) %>%
        full_join(wind_exp, by = c("storm_id", "fips")) %>%
        filter(as.numeric(gsub(".+-", "", storm_id)) >= 1996) %>%
        full_join(flood_exp, by = c("storm_id", "fips")) %>%
        full_join(tornado_exp, by = c("storm_id", "fips")) %>%
        mutate(dist_exposed = !is.na(dist_exposed),
               rain_exposed = !is.na(rain_exposed),
               wind_exposed = !is.na(wind_exposed),
               flood_exposed = !is.na(flood_exposed),
               tornado_exposed = !is.na(tornado_exposed)) %>%
        group_by(storm_id) %>%
        summarize(any_exposed = n(),
                  Distance = sum(dist_exposed),
                  Rain = sum(rain_exposed),
                  Wind = sum(wind_exposed),
                  Flood = sum(flood_exposed),
                  Tornado = sum(tornado_exposed)) %>%
        arrange(desc(any_exposed))

GGally::ggpairs(select(all_exp, -storm_id, -any_exposed),
                aes(alpha = 0.5)) + theme_few()
```

```{r county_number, echo = FALSE, results = "asis"}
a <- all_exp %>% 
        mutate(storm_id = paste0(gsub("-", " (", storm_id), ")")) %>%
        slice(1:20) %>%
        rename(Storm = storm_id, 
               "At least one metric" = any_exposed) %>%
        xtable(caption = "Top 10 storms in terms of numbers of counties assessed as `exposed` based on at least one of the exposure metrics, for storms between 1996 and 2011.",
                       label = "tab:county_number")
print(a, include.rownames = FALSE, floating = TRUE,
      table.placement = "p", caption.placement = "top", 
      booktabs = TRUE, comment = FALSE, timestamp = NULL)
```

We also mapped geographic patterns in the number of hits per county for distance, rain, and wind exposure when we incorporated the intensity of exposure for each county-storm pair (Figures \ref{fip:top_dist_map}, \ref{fip:top_rain_map}, \ref{fip:top_wind_map}); Tables \ref{tab:top_distance}, \ref{tab:top_rain}, \ref{tab:top_wind}). For these maps, we only plotted "hits" for the top 2,000 county-storm pairs (i.e., the 2,000 storm-county pairs with smallest distance to storm's closest approach, in the case of distance-based exposure, or the 2,000 storm-county pairs with highest precipitation, in the case of rain-based exposure).

```{r top_distance, echo = FALSE, results = "asis"}
a <- distance_exposure %>%
        arrange(storm_dist) %>% 
        slice(1:20) %>%
        left_join(county_centers, by = "fips") %>%
        select(storm_id, name, storm_dist) %>%
        mutate(storm_id = paste0(gsub("-", " (", storm_id), ")")) %>%
        rename(Storm = storm_id,
               County = name,
               "Distance (km)" = storm_dist) %>%
        xtable(caption = "Top storm-county pairs in terms of closest distance between county's population mean center and storm's closest approach.", 
               label = "tab:top_distance")
print(a, include.rownames = FALSE, floating = TRUE,
      table.placement = "p", caption.placement = "top", 
      booktabs = TRUE, comment = FALSE, timestamp = NULL)
```

```{r top_rain, echo = FALSE, results = "asis"}
a <- rain_exposure %>%
        arrange(desc(tot_precip)) %>% 
        slice(1:20) %>%
        left_join(county_centers, by = "fips") %>%
        select(storm_id, name, tot_precip) %>%
        mutate(storm_id = paste0(gsub("-", " (", storm_id), ")")) %>%
        rename(Storm = storm_id,
               County = name,
               "Rainfall (mm)" = tot_precip) %>%
        xtable(caption = "Top storm-county pairs in terms of most cumulative precipitation in the window from two days before to one day after a storm's closest approach to the county.", 
               label = "tab:top_rain")
print(a, include.rownames = FALSE, floating = TRUE,
      table.placement = "p", caption.placement = "top", 
      booktabs = TRUE, comment = FALSE, timestamp = NULL)
```

```{r top_wind, echo = FALSE, results = "asis"}
a <- wind_exposure %>%
        arrange(desc(max_sust)) %>% 
        slice(1:20) %>%
        left_join(county_centers, by = "fips") %>%
        select(storm_id, name, max_sust) %>%
        mutate(storm_id = paste0(gsub("-", " (", storm_id), ")"),
               max_sust = round(max_sust)) %>%
        rename(Storm = storm_id,
               County = name,
               "Maximum sustained wind (m / s)" = max_sust) %>%
        xtable(caption = "Top storm-county pairs in terms of most intense sustained winds.", 
               label = "tab:top_wind")
print(a, include.rownames = FALSE, floating = TRUE,
      table.placement = "p", caption.placement = "top", 
      booktabs = TRUE, comment = FALSE, timestamp = NULL)
```

```{r echo = FALSE, warning=FALSE, message=FALSE}
plot_top_x <- function(exposure_data, x = 2000, desc = TRUE,
                       exposure_var, eastern_states = eastern_states,
                       my_fips = my_fips, title){
        missing_fips <- my_fips[!(my_fips %in% exposure_data$fips)]
        missing_counties <- data.frame(fips = as.numeric(missing_fips),
                                       count = 0)

        exposure_data <- as.data.frame(exposure_data)
        exposure_data$exp_var <- exposure_data[ , exposure_var]
        if(!desc){ exposure_data$exp_var <- -1 * exposure_data$exp_var}

        exposure_x <- exposure_data %>%
                arrange(desc(exp_var)) %>%
                slice(1:x) %>% 
                group_by(fips) %>%
                summarise(count = n()) %>%
                rbind(missing_counties) %>%
                rename(region = fips) %>%
                mutate(region = as.numeric(region),
                       count = as.numeric(count))

        exposure_x$value <- cut(exposure_x$count,
                                c(0, 1, 3, 5, 7, 9, 11, 13, 100), 
                   labels = c("0", "1-2", "3-4", "5-6", "7-8",
                              "9-10", "11-12", "13+"), 
                   include.lowest = TRUE)
        exposure_palette <- c("#FFFFFF", brewer.pal(8, name = "YlGnBu"))

        map_top_x <- CountyChoropleth$new(exposure_x)
        map_top_x$set_zoom(eastern_states)
        map_top_x$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
                                                values = exposure_palette,
                                                na.value = "white")
        map_top_x$title <- title
        out <- map_top_x$render()
        return(out)
}
```

```{r top_dist_map, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Number of storms for which each county experienced one of the top 2,000 intensities for distance."}
plot_top_x(distance_exposure, desc = FALSE, exposure_var = "storm_dist",
           eastern_states = eastern_states, my_fips = my_fips,
           title = "Hits for top 2,000 Storm-County Pairs\nfor Distance Metric")
```

```{r top_rain_map, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Number of storms for which each county experienced one of the top 2,000 intensities for rain."}
plot_top_x(rain_exposure, exposure_var = "tot_precip",
           eastern_states = eastern_states, my_fips = my_fips,
           title = "Hits for top 2,000 Storm-County Pairs\nfor Rain Metric")
```

```{r top_wind_map, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Number of storms for which each county experienced one of the top 2,000 intensities for distance"}
plot_top_x(wind_exposure, exposure_var = "max_sust",
           eastern_states = eastern_states, my_fips = my_fips,
           title = "Hits for top 2,000 Storm-County Pairs\nfor Wind Metric")
```
